---
title: "Balaton 2019 Analysis"
author: "csaund"
date: "3/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(copBasic)
```

 Writeaforloopthatcreatesthevectorx=(3 5 7 9 1 1 1 3 1 5 1 7)′. Hint: Usethe formula n · 2 + 1. Employ two different approaches:
 
(a) Pre-allocate enough memory: Start with a zero vector of appropriate length and iteratively replace its elements. Hint: ?numeric
```{r}
x <- numeric(12)
x[1] = 3
x[2] = 5
# ...
```
(b) Grow an object: Start with a NULL object and iteratively append new results.
Verify with a larger vector and the system.time() function that the first approach
is more efficient.
```{r}
x2 <- NULL
x2 <- append(x2, 3)
x2 <- append(x2, 5)
# ...
```

b) Implement the following function in R:
 − 1 i f x < 0 , 
f(x) = 0 if x = 0, 1 ifx>0.
```{r}
pos_neg <- function(x) {
  if(x < 0) {
    return(-1)
  } else if(x == 0) {
    return(0)
  } else {
    return(1)
  }
}
```

c) Write a function that cumulatively adds the values of the vector x = (1,2,3,4,...,20)′. The result should look like y = (1,3,6,10,...,210)′. Solve this problem by
(a) using a for loop,
```{r}
X <- 1:20

cumulative_add <- function(x) {
  y <- NULL
  for(i in 1:length(x)) {
    x_sum = 0
    for(j in 1:i) {
      x_sum = x_sum + x[j]
    }
    y <- append(y, x_sum)
  }
  return(y)
}

y <- cumulative_add(X)
```

(b) matrix multiplication (vectorized solution),

(c) an R function (search for it with ??). Which solution is the most efficient?

(d) Write a function that implements some statistical procedure. This could either be
􏰀 a simple linear regression, yi = β0 + β1xi + εi; your function should estimate
the parameter vector β = (β0, β1)′ using, e. g., βˆ = (X′X)−1X′y; or
􏰀 a test for independence in a two-by-two contingency table that computes both
the Pearson-chi-square statistic
2 􏰂(xij −μˆij)2
X= μˆ , cells ij
 and the likelihood ratio statistic
G2 = 2 􏰂 xij log xij ,
where xij and μˆij = xi+x+j/x++ are the observed and expected frequencies, respectively; or
 1
cells μˆij
􏰀 any other statistical test that you find interesting (maybe search at Wikipedia for the formulae).
Assign the return value of your function a new class. Write a print method for this class.